---
title: Native advertising as a content marketing strategy
author:
  - name: Davor Trbusic
    email: davor.trbusic@unicath.hr
    affiliation: Croatian Catholic University
    correspondingauthor: true
    footnote: 1
  - name: Drazen Males
    email: dmales@fhs.hr
    affiliation: Faculty of Croatian Studies
  - name: Luka Sikic
    email: luka.sikic@unicath.hr
    affiliation: Croatian Catholic University
    footnote: 2
address:
  - code: Croatian Catholic University
    organization: Croatian Catholic University
    addressline: Ilica 242
    city: Zagreb
    state: 
    postcode: 10000
    country: Croatia
  - code: Faculty of Croatian Studies
    organization: Faculty of Croatian Studies
    addressline: Borongajska cesta 83d
    postcode: 10000
    city: Zagreb
    country: Croatia
footnote:
  - code: 1
    text: "This is the corresponding author."
  - code: 2
    text: "Co-author of the study."
abstract: |
  This research focuses on analyzing the content elements of native advertisements on the most popular Croatian online portals, with the aim of identifying key characteristics of headlines and content, as well as their correlation with audience reach. The study was conducted on a sample of 543 native ads published on six leading Croatian online portals from December 2021 to May 2022. The content analysis method was used, focusing on two general categories: headline characteristics and content characteristics. The analysis found that nearly 80% of the headlines had at least one clickbait characteristic, with the most commonly used styles including uncertainty (44.38%), use of numbers (9.39%), and emphasis on emotions (5.89%). The headlines were predominantly declarative sentences, with one-fifth being interrogative sentences, while the connection between the ad headline and the advertiser's brand name was very rarely present. In terms of content, visual elements were present in all ads, with photographs being the dominant element, and visual identity elements of the advertiser, such as logos, were included in 34.44% of the ads. Regarding sources in native ads, 47.33% of the ads did not use statements, while the most common sources were individuals from the organization (19.71%). Statements from direct users of the products or services were present in 6.81% of the cases, while statements from brand ambassadors and influencers were present in smaller percentages (4.79% and 3.5%, respectively). These findings highlight the dominance of clickbait headlines and the importance of visual identity in native advertising, while direct brand association and the use of various sources, including direct users, appear less frequently. A comparison of different types of headlines and sources with ad reach indicates specific practices in native advertising across different industries.
keywords: 
  - native advertising
  - content analysis
  - headline characteristics
journal: "An awesome journal"
date: "`r Sys.Date()`"
linenumbers: false
numbersections: true
bibliography: refs.bib
biblio-style: elsarticle-harv # author year style for natbib - use 'elsarticle-num' or 'elsarticle-num-names' for numbered scheme
classoption: preprint, 3p, authoryear # remove authoryear if not using `elsarticle-harv`
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
---


Please make sure that your manuscript follows the guidelines in the 
Guide for Authors of the relevant journal. It is not necessary to 
typeset your manuscript in exactly the same way as an article, 
unless you are submitting to a camera-ready copy (CRC) journal.

For detailed instructions regarding the elsevier article class, see   <https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions>

# Bibliography styles

Here are two sample references: @Feynman1963118 [@Dirac1953888].

By default, natbib will be used with the `authoryear` style, set in `classoption` variable in YAML and with `elsearticle-harv.bst` which is among provided style by `elsarticle` documentclass. Other available style are `elsarticle-num.bst` and `elsarticle-num-names.bst` — the first one can be used for the numbered scheme, second one for numbered with new options of natbib.sty. 

You can sets extra options with `natbiboptions` variable in YAML header. Example 
```yaml
natbiboptions: longnamesfirst,angle,semicolon
```

There are various more specific bibliography styles available at
<https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>. 
To use one of these, add it in the header using, for example, `biblio-style: model1-num-names`.

## Using CSL 

If `citation_package` is set to `default` in `elsevier_article()`, then pandoc is used for citations instead of `natbib`. In this case, the `csl` option is used to format the references. Alternative `csl` files are available from <https://www.zotero.org/styles?q=elsevier>. These can be downloaded
and stored locally, or the url can be used as in the example header.

# Equations

Here is an equation:
$$ 
  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
  \left(\frac{x}{\beta}\right)^{\alpha-1}
  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
  \alpha,\beta,x > 0 .
$$

Here is another:
\begin{align}
  a^2+b^2=c^2.
\end{align}

Inline equations: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$

# Figures and tables

Figure \ref{fig2} is generated using an R chunk.

```{r fig2, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig2}A meaningless scatterplot.", echo = FALSE}
plot(runif(25), runif(25))
```

# Tables coming from R

Tables can also be generated using R chunks, as shown in Table \ref{tab1} for example.

```{r tab1, echo = TRUE}
knitr::kable(head(mtcars)[,1:4], 
    caption = "\\label{tab1}Caption centered above table"
)
```

# References {-}

